---
title: <font color='grey' size='5em'>Prosper.com Data Set</font>
author: <font color='#3333ff' size='2em'>Tarique Hasheem</font>
date: <font color='#00FFFF' size='2em'>11/1/2017<font/>
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(knitr)
library(dplyr)
library(tibble)

```

<font color=grey>

## Load the Data

Prosper is an online platform where people can request to borrow money and 
investors can fund these requests.  Prosper manages the marketplace of both 
borrowers and investors, determines credit guidelines, and keeps a percentage of 
the revenue.  The data set contains 113,937 loans with 81 variables.

There are many data points to analyze however I'm interested in causes for 
delinquent and or charged off accounts?  Are there relationships between 
variables to explain these derogatory accounts?  If so, what are they and can 
they help mitigate future losses with more reliable borrowers?

For those who are curious to know what specific variables we can work with, I've 
taken the liberty to printing them out for your viewing pleasure alphabetically 
in table() format.

```{r echo=FALSE, warning=FALSE}
df <- read.csv("prosperLoanData.csv")
table(colnames(df))

```

## Univariate Plot Section

After quickly playing with the data set I realized running summary(), na.omit, 
sd(), var(), range(), and quantile() becomes real tedious so I created a 
function, sumFun(), which executes everything just mentioned in a few words of 
code.  If you care to see the construction of the function please reference the 
data Set.Rmd file.

#### Bar Plot of Loan Status

Lets look at a at bar plot of Loan Status to see if the visual gives us anything
we can sink out teeth into?

```{r echo=FALSE, warning=FALSE}
ggplot(df, aes(LoanStatus)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(limits = c(0, 70000), breaks = seq(0, 70000, 5000))

```

The plot does show us completed and current loans are more frequent than
derogatory items.  What's interesting is charge offs are more frequent than 
defaults.  I had imagined far greater defaults and less charge offs through loan 
work out's, lengthier terms, or forbearance agreements.  Good plot to get us
started.


#### Borrower APR

Let's begin with running our sumFun() function to better understand interest 
rates charged to borrowers.

```{r echo=FALSE, warning=FALSE}
# function code box
sumFun <- function(arg) {
  # function combines subset(df, !is.na()) and summary() into one
  col1 <- df[,c(arg)] # asign coloumn from datafram
  col1a <- na.omit(col1) # omit all NA values

  print(summary(col1a))
  print(paste("Standard Deviation", round(sd(col1a, na.rm = TRUE), 4)))
  print(paste("Variance", round(var(col1a, na.rm = TRUE), 4)))
  print(paste("Range", round(range(col1a, na.rm = TRUE), 4)))
  print("Quantile:")
  print(quantile(col1a, na.rm = TRUE))
  
}

```



Borrower APR Summary via sumFun() function.

```{r echo=FALSE, warning=FALSE}
sumFun('BorrowerAPR')

```


Okay, so we have a wide spectrum of interest rate charges.  Lets place the 
interest rates in a table and break them up by 0.5 % increments.

```{r echo=FALSE, warning=FALSE}
borrowerAPR <- df$BorrowerAPR # callable variable for our BorrowerAPR column
apr_breaks <- seq(.0065, .5200, by=.005) # breaking the range of interest rates by .005 increments or 0.5%
apr_cut <- cut(borrowerAPR, apr_breaks) # this will palce the interest rates in their respected breaks
apr_freq <- table(apr_cut) # lets use a table() format to print our results
apr_freq

```

The table helps but I want a visual to help me understand the distribution with a simple plot() function.  

```{r echo=FALSE, warning=FALSE}
plot(apr_freq)
```

Much better!  We can see there is a normal distribution where majority of interest rates are between 14% and 26%.  We can double check by running quantile() on its own.

```{r echo=FALSE, warning=FALSE}
quantile(borrowerAPR, na.rm = TRUE)
```

#### Lower Credit Score Range Histogram

Lets take a quick look at the distribution of the lower bound credit score ranges using a histogram plot.
```{r echo=FALSE, warning=FALSE}
ggplot(df, aes(CreditScoreRangeLower)) + 
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(500, 850))
```

We see that most of the borrowers have credit scores above 650 and we do have
a number of borrowers below 650 as well.

#### Upper Credit Score Range Histogram

```{r echo=FALSE, warning=FALSE}
ggplot(df, aes(CreditScoreRangeUpper)) + 
  geom_histogram(binwidth = 20) +
  scale_x_continuous(limits = c(500, 850))
```

The distribution is different than for lower bound credit scores but one thing is
for sure, prosper.com likes higher credit score borrower.


#### Borrower Rate

The difference between rate and APR is always part of due diligence for 
borrowers deciding on financing options.  Even though in some situations APR 
makes decisions extremely confusing.  Its always a good idea to understand if 
the APR amortizes the cost associated with the loan throughout the entire term 
or just the first year?  Simply put, fees for originating loans are only 
collected once, at time of origination and they never reoccur.  In either event 
its a good idea for us to understand how the consumer see's their cost.  Lets 
look at the actual rate borrowers are quoted.

Lets begin with a summary.

```{r echo=FALSE, warning=FALSE}
borrowerRate <- df$BorrowerRate

sumFun("BorrowerRate")

```

No surprise here, the mean borrower rate is .1928 or 19.28% and the mean 
borrower APR is 21.88%, a mean difference of 2.6%.  Now lets  look at a more 
colorful histogram plot of borrower rate.

```{r echo=FALSE, warning=FALSE}
borrowerRate <- ggplot(df, aes(BorrowerRate))
borrowerRate + geom_histogram(fill = 'green', color = 'purple', binwidth = .01) +
  scale_x_continuous(limits = c(.05, .40), breaks = seq(.05, .40, .05)) +
  scale_y_continuous(limits = c(0, 7000), breaks = seq(0, 7000, 500))
```

 
#### Lender Yield

What good is interest if there's no yield.  Assuming Prosper is always on-top of 
usury laws lets understand the amount of return or yields investors/ lenders can 
anticipate?  Can this variable be used later in our analysis to paint a picture 
for delinquent or charge off accounts?

```{r echo=FALSE, warning=FALSE}
sumFun('LenderYield')
```

It's obvious lender yields will correlate with borrower APR and rate.  And the 
summary does not really tell us any more than we already knew, so for now we'll 
leave this variable on the back burner.  Lets still plot it for us to have a 
visual.

```{r echo=FALSE, warning=FALSE}
lenderYield <- ggplot(df, aes(LenderYield))
lenderYield + geom_histogram(fill = '#9900cc', color = '#13001a', binwidth = .01) +
  scale_x_continuous(limits = c(.05, .40), breaks = seq(.05, .40, .05)) +
  scale_y_continuous(limits = c(0, 7000), breaks = seq(0, 7000, 500))
```

#### Estimated Loss

We all know derogatory accounts will show up as a loss somewhere or another so 
lets run a summary on estimated loss to see what's in store for us?

Estimated Loss Summary

```{r echo=FALSE, warning=FALSE}
sumFun("EstimatedLoss")
```

The problem here, and I should have caught this earlier is that we are not sure 
how these losses are calculated, in other words the mean estimated loss is 
0.08031 but we're not sure if this is from total revenue, from lender yield, 
from gross, or what?  This variable may come in handy later on but for now its 
simply to vague. 

Estimated Loss Histogram Plot

```{r echo=FALSE, warning=FALSE}
estimatedLoss <- ggplot(df, aes(EstimatedLoss))
estimatedLoss + geom_histogram(fill = "#00ffff", color = "#004d4d", binwidth = .01) +
  scale_x_continuous(limits = c(.004, .40), breaks = seq(.004, .40, .05)) +
  scale_y_continuous(limits = c(0, 7000), breaks = seq(0, 7000, 500))

```

#### Loan Origination Amount

The original loan amount will surely help us navigate our data set.  We've all 
been there before, borrowing more than we’re comfortable with.  Will loan 
amounts have an impact on understanding delinquencies?  Lets begin with running 
a summary be ready to combine this variable with others in our exploration to 
see what role loan amount play? 

Loan Origination Amount Summary

```{r echo=FALSE, warning=FALSE}
sumFun("LoanOriginalAmount")
```

The loan amounts are reasonable amounts for borrowers considering debt 
consolidation.  When reviewing Prosper.com this is a main reason for borrowers 
seeking funding.  

Loan Original Amount Histogram Plot

```{r echo=FALSE, warning=FALSE}
loanOriginal <- ggplot(df, aes(LoanOriginalAmount))
loanOriginal + geom_histogram(fill = "#0099ff", color = "#004d4d", binwidth = 500) +
  scale_x_continuous(limits = c(1000, 36000), breaks = seq(1000, 36000, 3000))

```

Its not the best looking distribution but it gives us an idea of the direction 
of the tail.


#### Debt to Income Ratio

Debt to income ratio is a key variable when creditors and lenders decide on an 
applicants request for funding.  Lets take a look at our summary.

```{r echo=FALSE, warning=FALSE}
sumFun("DebtToIncomeRatio")
```

We see we have outliers of 10.0100, there must be some sort of glitch, if we 
move our decimal to the right two places these borrowers have 1,000% debt to 
income ratio, very unlikely.  But why is it happening?

Debt to Income Ratio Histogram Plot

```{r echo=FALSE, warning=FALSE}
debt2income <- ggplot(df, aes(DebtToIncomeRatio))
debt2income + geom_histogram(fill = '#339933', color = '#13001a', binwidth = 0.1) +
  scale_x_continuous(limits = c(.1, 1), breaks = seq(.1, 1, .05)) 
  #scale_y_continuous(limits = c(0, 7000), breaks = seq(0, 7000, 500))
```

When we cut away the outliers the graph look reasonable.  More money loans 
originated to borrowers with lower debt to income ratio's that higher ratios.

## Univariate Analysis

#### What are the structure of your data set?

There are 113.9 thousand row’s of data and 81 features (variables).  

Preliminary observations:

* 21.88% - mean borrower APR
* 19.28% - mean borrower rate
* 18.27% - mean lender yield
* 08.03% - mean estimated loss
* $8,337 - mean original loan amount

#### What are the main features of interest in your data set?

The data set is fascinating and we can subset() and or select() so many 
variables and dive deeper into our data set.  For now I'll lets focus on loan 
status in particular derogatory accounts.

#### What other features in the data set do you think will help support your 
#### investigation into your features of interest?

* Loan Status
* Occupation
* Revolving Credit Balance
* Bank Card Utilization
* Income Verifiable
* Credit Score Range (Lower, Upper)


#### Did you create any new variables from existing variables in the data set?

Yes, and to be sure we're on the same page, below we have a new variable as the 
frequency distribution for the LoanStatus variable.  My goal here is to better 
understand the frequency of derogatory items such as charge offs and defaulted.

```{r echo=FALSE, warning=FALSE}
loanStatus <- df$LoanStatus # loanStatus column variable
loanStatus.freq = table(loanStatus) # applies the table function

# cbind will print the results in a column format
cbind(loanStatus.freq)

```

#### Of the features investigated, were there any unusual distributions?  
#### Did you perform any operations?  If so, why?

The debt to income variable is interesting, there are over 200 debt to income 
ratio of 10.  In other words the borrowers debt is 10 times greater than their 
income.  This is unusually high.  The operation I performed was to better 
understand the frequency distribution.

## Bivariate Plots Section

#### Loan Status vs Borrower APR

Great, now lets look at two variables at the same time.  We'll begin with loan 
status and borrower APR.  As we know from above loan status a categorical 
variable and will play nicely with borrower APR when we plot.  Lets also print
outliers for us to review.

```{r echo=FALSE, warning=FALSE}
# the funciton will return a boolean TRUE/FALSE for all values ina a variable
# to determine outliers.

is_outlier <- function(x) {
  return(x < quantile(x, 0.25, na.rm = TRUE) - 1.5 * IQR(x, na.rm = TRUE) | 
           x > quantile(x, 0.75, na.rm = TRUE) + 1.5 * IQR(x, na.rm = TRUE))
}

# we create new variable dat
dat <- df %>% 
  # we create a new coloumn named 'outlier' 
  tibble::rownames_to_column(var="outlier") %>% 
  # we group_by a variable we are interested in
  group_by(LoanStatus) %>% 
  # mutate will calcuate new values for our new column 'outliers'
  # it will run the is_outlier function from above on a variable from the dataframe
  mutate(is_outlier=ifelse(is_outlier(BorrowerAPR), BorrowerAPR, as.numeric(NA)))

# filtering the oulier variable
dat$outlier[which(is.na(dat$is_outlier))] <- as.numeric(NA)

# plot the new variable 'dat' and print out outliers on the plot
ggplot(dat, aes(BorrowerAPR, LoanStatus)) +
  geom_boxplot(color = '#00FFFF', fill = 'blue', alpha=0.2, notch = TRUE,
               notchwidth = 0.9, outliers.colour='red', outlier.fill='red',
               outlier.size = 3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_text(aes(label=is_outlier),na.rm=TRUE,nudge_y=0.05)
```

It's a difficult plot to make out.  Lets reverses our axes for a better looking 
plot. 

```{r echo=FALSE, warning=FALSE}
ggplot(dat, aes(LoanStatus, BorrowerAPR)) +
  geom_boxplot(color = '#7FFFD4', fill = '#008080', alpha=0.7, notch = TRUE,
               notchwidth = 0.9, outliers.colour='red', outlier.fill='red',
               outlier.size = 3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_text(aes(label=is_outlier),na.rm=TRUE,nudge_y=0.05)
```

Much better.  We see the different loan status and outliers but still not a 
clear picture.

#### Credit Grade vs Borrower APR

Lets see if credit grade has any affect on Borrower APR?  We'll print the 
outliers as well.

```{r echo=FALSE, warning=FALSE}
dat2 <- df %>% 
  tibble::rownames_to_column(var="outlier") %>% 
  group_by(CreditGrade) %>% 
  mutate(is_outlier=ifelse(is_outlier(BorrowerAPR), BorrowerAPR, as.numeric(NA)))

dat2$outlier[which(is.na(dat$is_outlier))] <- as.numeric(NA)

ggplot(dat2, aes(CreditGrade, BorrowerAPR)) +
  geom_boxplot(color = '#00cc00') +
  geom_text(aes(label=is_outlier),na.rm=TRUE,nudge_y=0.05, 
            check_overlap = TRUE, size = 3) +
  scale_y_continuous(limits = c(0, .55), breaks = seq(0, .55, .05)) 


```

Credit grade 'A' is peculiar.  After carefully reviewing the graph is seems that 
AA is a higher credit grade than A however due to logic AA is alphabetically 
graphed after A.  In other-words they are switched around and if you consider 
the switch it seems lower credit grades have higher interest rates.


#### Total Credit Lines Past 7 Years vs Loan Status

Lets take a closer look at trade lines and loan status.  Can having more trade 
lines affect delinquency?

```{r echo=FALSE, warning=FALSE}

dat3 <- df %>% 
  tibble::rownames_to_column(var="outlier") %>% 
  group_by(LoanStatus) %>% 
  mutate(is_outlier=ifelse(is_outlier(TotalCreditLinespast7years), TotalCreditLinespast7years, as.numeric(NA)))

dat3$outlier[which(is.na(dat$is_outlier))] <- as.numeric(NA)

ggplot(dat3, aes(LoanStatus, TotalCreditLinespast7years)) +
  geom_boxplot(color = '#ff3300', fill = '#66ff99', notch = FALSE) +
  geom_text(aes(label=is_outlier),na.rm=TRUE,nudge_y=0.05, 
            check_overlap = TRUE, size = 3) +
  scale_y_continuous(limits = c(0, 150), breaks = seq(0, 150, 10)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```

There might be something here.  It seems that the greater the amount of credit 
trade lines the less borrowers are likely to default, once borrowers get above 
75 trade lines.  This is good progress towards our analysis.

#### Total Trades vs Loan Status

```{r echo=FALSE, warning=FALSE}
dat4 <- df %>% 
  tibble::rownames_to_column(var="outlier") %>% 
  group_by(LoanStatus) %>% 
  mutate(is_outlier=ifelse(is_outlier(TotalTrades), TotalTrades, as.numeric(NA)))

dat4$outlier[which(is.na(dat$is_outlier))] <- as.numeric(NA)

ggplot(dat4, aes(LoanStatus, TotalTrades)) +
  geom_boxplot(color = '#ff00ff', fill = '#ffff00', alpha=0.7, notch = TRUE,
               notchwidth = 0.9, outlier.size = 1) +
  geom_text(aes(label=is_outlier),na.rm=TRUE,nudge_y=0.05, 
            check_overlap = TRUE, size = 3) +
  scale_y_continuous(limits = c(0, 150), breaks = seq(0, 150, 10)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

The graph above is even more promising for borrowers who historically have large 
number of trade lines without limiting for number of years as the previous graph 
showed.

#### Revolving Credit Balance vs Loan Status



Lets see if revolving credit card balances plays a role in defaults?

```{r echo=FALSE, warning=FALSE}
 ggplot(aes(LoanStatus, 
            RevolvingCreditBalance), 
        data = df) +
  geom_jitter( alpha = .3)  + 
  geom_boxplot( alpha = .5,color = 'blue')+ 
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)  +scale_y_log10()+
      theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Borrowers with greater revolving balances tend to keep their good standings.  
We'll keep these variables into consideration as we move along.

## Bivariate Analysis

#### Talk about some of the relationships.  How did the features of interest 
#### vary with other features in the dataset?

My goal is to better understand what variables or patterns play a role in 
understanding borrower defaults.  I began with plotting 'LoanStatus' with 
'BorrowerAPR' to see if we can determine if certain groups of interest rates 
have greater defaults when compared to others.  On the second plot I kept the 
variable 'BorrowerAPR' and but added 'CreditGrade' to see if there is some sort 
of logic behind credit rating and APR.  Then I plotted 'LoanStatus' with 
'TotalCreditLinesPast7years' and 'TotalTrades'.  The difference caps trade-lines 
at 7 years and the later is a historical variable without limiting for specific 
number of years.  The last plot keeps 'LoanStatus' and plots 
'RevolvingCreditBalance'.  

These features are different than others in the data set because they are more 
granular in personal detail that make up a borrowers credit history which is a 
direct reflection on interest rates. 

#### Did you observe any interesting relatinships between the other features?

The first couple of plots did not give us a whole lot to be excited about.   
However as we plotted ‘LoanStatus’ against ‘TotalCreditLines’ there are reasons 
to believe trade lines do play a role in default rates.  The greater the number 
of trade-lines the least likely a borrower defaults.  This is especially the 
case with borrower with 75 or greater historical trade-lines.

#### What was the strongest relationship you found?

The strongest relationship I found was between 'TotalTrades' and 'LoanStatus'.  
It looks as if historically borrowers with greater than ~60 or so trade-lines 
tend not to default when compared with borrowers with fewer trade lines. 

## Multivariate Analysis

Lets clear up some of the clutter and filter() 'Defaulted' | 'Chargedoff' 
from our LoanStatus variable. 

```{r echo=FALSE, warning=FALSE}
# the filter will replace defaults <- df$LoanStatus == 'Defaulted' and defaults <- df[defaults,]
filter_Defaulted.and.Chargedoff <- filter(df, LoanStatus == 'Defaulted' | LoanStatus == 'Chargedoff')
head(filter_Defaulted.and.Chargedoff)
```

The new variable above looks great.  Now lets plot CreditScoreRangeUpper, 
StateMonthlyIncome, and LoanStatus.

```{r echo=FALSE, warning=FALSE}
df$CreditScoreRangeUpper.bucket = cut(df$CreditScoreRangeUpper,
                            c(0, 500, 600, 700, 800, 900))

ggplot(aes(x =  (CreditScoreRangeUpper.bucket),
           y = StatedMonthlyIncome ), 
           data =subset( df, LoanStatus=='Defaulted'| LoanStatus== 'Chargedoff')) + 
   geom_boxplot( aes(fill= LoanStatus))  +
  scale_fill_brewer(type='seq',
                   guide=guide_legend(title='Quality'))  +
  ylim(0,2500)+
          theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

It's a clean plot but it doesn't tell us much.  Let's use the dplyr package to 
group our new variable by 'LoanStatus', 'StatedMonthlyIncome', 'TotalTrades', 
'CreditScoreRangeUpper' and we'll calculate the mean and median for 
'BorrowerAPR' into a new variable.  Lets print out a few rows to make sure all 
is well.

```{r echo=FALSE, warning=FALSE}
groupBy_filter_Defaulted.and.Chargeoff <- filter_Defaulted.and.Chargedoff %>%
  group_by(LoanStatus, StatedMonthlyIncome, TotalTrades, CreditScoreRangeUpper) %>%
  summarise(BorrowerAPR.mean = mean(BorrowerAPR),
            BorrowerAPR.median = median(BorrowerAPR),
            n = n())

head(groupBy_filter_Defaulted.and.Chargeoff)
```

Lets lot the new variable and see if we can build on our assumption from earlier 
that borrowers with less than 60 trade lines are more likely to have defaults 
and charge offs.  We'll run both log10 and coord_trans to see the visual 
difference in our plots.

```{r echo=FALSE, warning=FALSE}
ggplot(groupBy_filter_Defaulted.and.Chargeoff, 
       aes(BorrowerAPR.mean, TotalTrades, color = LoanStatus)) +
  geom_point(alpha = 1 / 5) +
  scale_y_log10()
  
```

Our coord_trans plot.

```{r echo=FALSE, warning=FALSE}
ggplot(groupBy_filter_Defaulted.and.Chargeoff, 
       aes(BorrowerAPR.mean, TotalTrades, color = LoanStatus)) +
  geom_point(alpha = 1 / 5) +
  coord_trans(y = 'sqrt') 
```

Here is the same three variables but with an added elipsed layer.

```{r echo=FALSE, warning=FALSE}

ggplot(aes(BorrowerAPR, TotalTrades, color = LoanStatus),
       data =subset( df, LoanStatus=='Defaulted'| LoanStatus== 'Chargedoff') ) +
  geom_point(alpha=0.1) +
  stat_ellipse(size=1)
```



Now lets plot our new variable but change the our x-axis to 
'CreditScoreRangeUpper'.

```{r echo=FALSE, warning=FALSE}
ggplot(groupBy_filter_Defaulted.and.Chargeoff, 
       aes(CreditScoreRangeUpper, TotalTrades, color = LoanStatus)) +
  geom_point() +
  scale_x_continuous(limits = c(550, 900), breaks = seq(550, 900, 50))
```

The plot above seems over plotted, lets revise is below.

```{r echo=FALSE, warning=FALSE}
ggplot(subset(df,LoanStatus == 'Defaulted' | LoanStatus == 'Chargedoff'), 
       aes(factor(CreditScoreRangeUpper), TotalTrades, fill = LoanStatus)) +
  geom_boxplot() +
          theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ylim(0,60)
```

Much better.  We see that borrowers above 60 or so trade lines have fewer
delinquencies or charge offs.

#### BorrowerAPR, CreditGrade, LoanStatu

Lets take a slight detour here and look at box plots.  Here we'll plot Credit
Grade, BorrowerAPR, and LoanStatus.


```{r}
dat4 <- df %>% 
  tibble::rownames_to_column(var="outlier") %>% 
  group_by(CreditGrade) %>% 
  mutate(is_outlier=ifelse(is_outlier(BorrowerAPR), BorrowerAPR, as.numeric(NA)))

dat2$outlier[which(is.na(dat$is_outlier))] <- as.numeric(NA)

ggplot(aes(CreditGrade, BorrowerAPR), data = subset(dat4, LoanStatus == 'Defaulted' | LoanStatus == 'Chargedoff')) +
  geom_boxplot(color = '#00cc00') +
  geom_text(aes(label=is_outlier, colour = LoanStatus),na.rm=TRUE,nudge_y=0.05, 
            check_overlap = TRUE, size = 3) +
  scale_y_continuous(limits = c(0, .55), breaks = seq(0, .55, .05)) 
```

The plot is a little crowded.  Lets see if we can revise it.

```{r}
ggplot(aes(ProsperRating..Alpha., BorrowerAPR, fill=LoanStatus), data = 
         subset(df, LoanStatus == 'Defaulted' | LoanStatus == 'Chargedoff')) +
  geom_boxplot() +
  scale_y_continuous(limits = c(0, .55), breaks = seq(0, .55, .05))
```

That's nice.  We see where our outliers are within each credit grade tier.


#### Talk about some of the relationships you observed in this part of the \ 
#### investigation. Were there features that strengthened each other in terms 
#### of \ looking at your feature(s) of interest?

I created a new variable using the filter() method to clear up the clutter from 
the original data frame and in particular the ‘LoanStatus’ variable.  This 
variable had 12 different categories so it was a good idea.  Our new filtered 
variable was still too cumbersome so I took advantage of the group_by() method 
in dplyr to only look at LoanStatus, StatedMonthlyIncome, TotalTrades, 
CreditScoreRangeUpper and summarized the mean and median for BorrowerAPR.  The 
importance of this process was it allowed me to isolate the features explicitly 
to look and allowed a more granular look at the remainder of my analysis.   

### Were there any interesting or surprising interactions between features?

Yes there are a number of surprising interactions.  We see that at an alpha of 
1/10 there are distinct vertical lines in the BorrwerAPR.mean axis.  In other 
words there are noticeable amounts of defaulted and charged off loans at 
approximately .30, .36 and .37 BorrowerAPR.mean and interestingly enough these 
lines fade once borrowers reach 60 trade lines.  

## Final Plot and Summary

#### Plot One

```{r echo=FALSE, warning=FALSE}

ggplot(groupBy_filter_Defaulted.and.Chargeoff,
       aes(BorrowerAPR.mean, TotalTrades, color = LoanStatus)) +
  geom_point(alpha = 1 / 10) +
  coord_trans(y = 'sqrt') +
  scale_x_continuous(breaks = seq(.05, .50, .05)) +
  scale_y_continuous(limit = c(0, 90), breaks = seq(0, 90, 10)) +
  labs(color = 'Loan Status', y = 'Number of Trade Lines', 
       x = 'Mean of Borrower APR in %', 
       title = 'Loan Status in Relation to Borrower APR and Number of Trade Lines') +
  stat_ellipse(size=1)
  
  
```


#### Description

The reason why I choose the plot above is because it shows two interesting finds.  
First, its shows that borrowers with 60 or greater trade lines are less likely 
to go into default or charge off status.  Second, we notice three defined 
vertical lines that suggest borrowers paying above 30% APR are more likely to 
go into default and charge off status.  

#### Plot two


```{r}
ggplot(subset(df,LoanStatus == 'Defaulted' | LoanStatus == 'Chargedoff'), 
       aes(factor(CreditScoreRangeUpper), TotalTrades, fill = LoanStatus)) +
  geom_boxplot() +
          theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ylim(0,60) +
    labs(color = 'Loan Status', y = 'Number of Historic Trade Lines Open', 
       x = 'Upper Credit Score Range', 
       title = 'Loan Status in Relation to Loan Status and Trade Lines')
```


#### Description

We see our median trade lines are about 30 and outliers in the 60 range.  
We also notice defaults and charge offs are common among all credit profiles.  
What is interested and to our point is that there are not a whole lot of 
outliers above 60 therefore borrowers with 60 plus trade lines would seem to 
be less riskier.


#### Plot three

```{r echo=FALSE, warning=FALSE}

lenderYield <- ggplot(df, aes(LenderYield))
lenderYield + geom_histogram(fill = '#9900cc', color = '#13001a', binwidth = .01) +
  scale_x_continuous(limits = c(.05, .40), breaks = seq(.05, .40, .05)) +
  scale_y_continuous(limits = c(0, 7000), breaks = seq(0, 7000, 500)) +
  labs(x="Lender Yield %", y="Number of Loans Originated", 
       title = "Lender Yield in Relation to Number of Loans Originated")
```

#### Description

Whats interested about the plot above is that we see a spike of number of loans 
originated where the lender yield is slightly above 30%.  As we saw in plot one 
These are the borrowers who visually have more defaults and charge offs.  

## Refelction

#### What were some of the struggles that you went through?

Plotting on the correct axes was a challenge at first especially when 
categorical and continuous variables were being plotted.   I also did not enjoy 
running summaries and different numerical measures such as range(), then I 
created the sumFun() function and this helped repeating the same lines of code 
over and over. 

#### What went well?

Describing the plots went well.  Labeling the axes, title, and legends was fun 
and rewarding knowing the plot was finally complete.

#### What was surprising?

What was surprising borrowers who have greater than 60 trade lines have fewer 
defaults and charge offs.  Its a surprise because credit score was not much of 
a factor here.  People on both ends of the credit spectrum (high and low scores)
performed better when they had 60 or more trade lines.

#### Insight on Future Work?

The data set is rich with variables.  You can spend months analyzing different 
variables and their relation with one another.  I’d be interested in diving 
deeper to sort through each variable in relation to loan status.  I’m sure there
are more variables other than trade lines that can help us better understand 
borrower patterns into default and charge offs.

<br><br>
<br><br>
<br><br>
<br><br>

</font>